Project 1 AS 2014-2015: birthweight as fuction of parity
==============================================================================
  
  <big>
  **Daphné RIMSKY-ROBERT and Pauline SMITH**  
  **27/01/15**
  </big>
  
```{r setup, include=FALSE}
opts_knit$set(progress=FALSE, verbose=FALSE, width=90)
opts_chunk$set(message=FALSE, tidy=TRUE, highlight=TRUE, fig.align="center")
```


library(knitr)
library(markdown)
knit("Markdown.Rmd", quiet=TRUE)
markdownToHTML("Markdown.md", "Markdown.html", option=c("highlight_code", "toc"), title="Simple R Markdown template")
browseURL("Markdown.html")


<h1> Importing data </h1>

```{r load, results='hide'}
library(foreign)
library(lattice)
library(psych)
wd = "C:/Users/SexyManatee/Documents/GitHub/Rstat_projet"
w <- read.spss(paste(wd, "01-weights.sav", sep="/"), to.data.frame=TRUE)
names(w)
str(w)
```

```{r mask, echo=FALSE}
library(xtable)  ## useful for displaying HTML table
```

<h1> Question 1 </h1>

### Numerical summary of sample size (frequencies and counts) according to parity

Creating functions to process data:
```{r functions}
#counts
func_count = function(x, digits = 1)
  round(length(x), digits = digits)

#frequencies
func_freq = function(x, digits = 1)
  round(length(x)/length(w$GENDER), digits = digits)

```{r variables, results='asis'}

```


Processing data:
```{r variables_countfreq, results='hide'}

#counts
d_count = aggregate(w$GENDER ~ PARITY, data = w, func_count)
names(d_count) = c("PARITY", "COUNT")

#frequencies
d_freq = aggregate(w$GENDER ~ PARITY, data = w, func_freq)
names(d_freq) = c("PARITY", "FREQUENCY")
```


Printing table for frequencies and counts:
```{r xtable_countfreq, results='asis'}

countfreq_summary = data.frame(d_count, d_freq[2])
print(xtable(countfreq_summary), type = "html", include.rownames = TRUE)

```




### Numerical summary of baby weights according to parity

Creating function:
```{r function_weight}
func_weight = function(x, digits = 1)
  round(c(mean = mean(x), sd = sd(x), range = range(x)), digits = digits)
```

Processing data:
```{r variables_weight, results='hide'}
d_weight = aggregate(WEIGHT ~ PARITY, data = w, func_weight)
d_weight = as.data.frame(as.matrix(d_weight))
names(d_weight) = c("PARITY", "MEAN", "STAND DEV", "RANGE MIN","RANGE MAX")
```

Printing table for baby weights:
```{r xtable_weight, results='asis'}
print(xtable(d_weight), type = "html", include.rownames = TRUE)
```

Density plot:
```{r densityplot, results='asis'}
densityplot(~WEIGHT, data = w, groups = PARITY, auto.key = TRUE)
```


<h1> Question 2 </h1>

### ONE-WAY ANOVA
```{r aov_4_levels}
mod = aov(WEIGHT~PARITY, data = w)
print(summary(mod), type = "html", include.rownames = FALSE)
```

p < 0.03 : we reject the null hypothesis and conclude that there is a link between parity and baby weights.

### part of the variance explained
It is computed using the following formula:
 SS(effect)/ss(effect + residuals)
```{r part_of_var_4levels, results='asis'}
part_explained = summary(mod)[[1]][["Sum Sq"]][1]/summary(mod)[[1]][["Sum Sq"]][2]
print(part_explained, type = "html", include.rownames = FALSE)
```
The part of the variance explained here is about 1.8%. 
### Assumptions
<ul>
**<li><p>Independance</p>**
<p>The sample includes less than 10% of the population and no obvious links are reported between the children. We can assume that the samples are independant.</p></li>
**<li><p>Approximately normal</p>**
<p>We need to compute a qqplot:</p></li>
</ul>
```{r qqnorm_plot, results='asis'}
qqnorm(w$WEIGHT)
```
It looks like a roughly linear function relates our sample distribution and the normal distribution, thus we can conclude that the sample distribution is approximately normal.

<ul>
**<li>Constant variance</li>**
</ul>
```{r}
bwplot(PARITY ~ WEIGHT, data = w, pch = "|")
```
According to what can be observed hereand the data constained in the tables of question 1, it seems that the width of the boxes for each group is relatively constant, though "3 or more siblings" is slightly off.

<h1> Question 3 </h1>

### Aggregating the last two categories:
```{r aggregate}
levels(w$PARITY)[4] = levels(w$PARITY)[3]
levels(w$PARITY)[3] = "2 or more siblings"
```


### ONE-WAY ANOVA on new groups
```{r aov_3levels}
mod_2 = aov(WEIGHT~PARITY, data = w)
print(summary(mod_2), type = "html", include.rownames = FALSE)
```
p< 0.01: we still reject the null hypothesis and conclude there is a link between parity and baby weights.


### part of variance explained
```{r pov_3levels}
part_explained_2 = summary(mod_2)[[1]][["Sum Sq"]][1]/summary(mod_2)[[1]][["Sum Sq"]][2]
print(part_explained_2,type = "html", include.rownames = FALSE)
```
The part of the variance explained here is still about 1.8%.
### Assumptions
<ul>
**<li><p>Independance</p>**
<p>No changes in this assumption</p></li>
**<li><p>Approximately normal</p>**
<p>We need to compute a qqplot:</p></li>
</ul>
```{r qqnorm, results='asis'}
qqnorm(w$WEIGHT)
```
No changes in this assumption
<ul>
**<li>Constant variance</li>**
</ul>
```{r}
bwplot(PARITY ~ WEIGHT, data = w, pch = "|")
```
According to what can be observed here and the data constained in the tables of question 1, it seems that the width of the boxes for each group is fairly constant. The assumption of constant variance is slightly better respected after aggregating the last two categories.

<h> Question 4 </h>

### Linear regression
```{r}
m = lm(w$WEIGHT ~ w$PARITY)
print(summary(m), type = "html", include.rownames = FALSE)
```

p < 0.01: we reject the null hypothesis and conclude that there is a linear correlation between baby weight and parity. 

**R-squared** is a measure of the effect-size.
```{r rsq}
print(summary(m)$r.squared, type = "html", include.rownames = FALSE)
```
R-squared is close to 0, we can conclude that the correlation is low, even if present.
